Simulated observations of young protoplanetary discs

### 2012-12-19 paper.tex
now using eps rather than png

\title[Young protoplanetary discs]{Simulated observations of young protoplanetary discs}
\author[Tom  Douglas, Paola Caselli, Et al.]{Tom Douglas$^{1}$\thanks{E-mail:
pytd@leeds.ac.uk}, Paola Caselli$^{1}$, Aaron Boley$^{2}$, Richard Durisen$^{3}$,\newline Tom Hartquist$^{1}$, John Ilee$^{1}$, Jonathon Rawlings$^{4}$ \\
$^{1}$School of Physics and Astronomy, University of Leeds, Leeds LS2 9JT, UK \\
$^{2}$Department of Astronomy, University of Florida, 211 Bryant Space Center, PO Box 112055, USA\\
$^{3}$Department of Astronomy, Indiana University, 727 East 3rd Street, Swain West 319, Bloomington, IN 47405, USA\\
%$^{4}$Department of Physics & Astronomy, University College London, London WC1E 6BT, UK\\
}
\begin{document}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/L1544model_used_legend.png}
 \label{fig:l1544_model}
 \caption{[NOTE: YOU NEED TO ADD LABELS AND UNITS TO THE AXES!!] Model of the pre-stellar core L1544 used as the envelope of the young protoplanetary disc in the hybrid model. Showing temperature (red) and dust (green) temperature in kelvin, log number density (blue) in cm$^{-3}$ and inward velocity$\times$100 (cyan) in m$\,$s$^{-1}$. Adapted from Keto \& Caselli (2010) and Keto et al. (2012, in preparation).}
\end{figure}

@@

\begin{figure*}
 \includegraphics[width=168mm]{Figures/model/columnDensities2.png}
 \label{Chemistry} 
 \caption{Column densities of OCS, H$_2$CO and HCO$^+$ used in the disc model. Figure adapted from Ilee at al. (2011) [SHOULD WE ALSO ADD C18O COLUMN DENSITY AS WE WANT TO DISCUSS THAT LATER ?? ]}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/rotational_velocities.png}
 \label{velocity}
 \caption{Azimuthaly averaged rotational velocity in the disc mid-plane with best fit curve}
\end{figure}

@@

\subsection{The radiative transfer code} \label{subsec:radiative_transfer_code}
%-describe the RT used (LIME) 
The radiative transfer program used, LIME (Brinch \&Hogerheijde 2010), calculates line intensities based on a weighted sample of randomly chosen points in a continuous 3D model. The method of selecting these points is given in the grid-construction section. At each of these points, the density of the main collision partner (H$_2$), gas and dust temperatures, velocity, molecular abundances and turbulent velocity are taken from the model. These points are then smoothed by Lloyds algorithm (Lloyd 1982) in order to minimise the variation in distance between points whilst keeping the same underlying distribution. These points are then connected by Delaunay triangulation \footnote{In three dimensions this means that if four points are conneted into a tetrahedron, the sphere circumscribing these four points contains no other points. It can be shown that this connection is unique for a given set of points.} and it is down these paths that photon propagation is restricted (figures \ref{grid}). The level populations of the selected molecules are calculated at each of these points from collisional and radiative (de)excitation and the local radiation field is calculated. This is repeated 20 times with the populations of each level converging towards a single value. This number of iterations is sufficient for the signal to noise ratio of the level populations (as defined in Brinch \&Hogerheijde 2010) to exceed 1000 in 99\% of the points. [NOTE: WHY DO WE NEED TO REACH SUCH A HIGH S/N? --- it isn't s/n in the usual sense, it's a measurement of how converged the model is, defined in brinch 2010 as change in populations this iteration / std-dev of change in pops over last 5 iterations] \newline
After 20 iterations of updateing the level populations the model is ray-traced in order to produce synthetic brighness maps. In order to minimise the ``blockyness'' of the output images, resulting from the grid construction, the average of ten seperate runs was taken (figure \ref{averages}).

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/average_comparison.png}
 \label{averages}
 \caption{A plot showing the difference between a single LIME continuum image at 300$\,$GHz (right) and the average of ten such images from different runs of the same model in LIME (left).}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/Lime_grid3.png}
 \label{grid}
 \caption{A plot of the points selected by the griding process and the paths down which photons can propagate overlaid on a smoothed density model. The points are more concentrated at small radii and in the densest regions.}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/lime_points_rz_histo.png}
 \label{points}
 \caption{A 2D histogram of the point distribution throughout the model. The disc and envelope can be seen as two separate entities which have to be sampled using different point distributions}
\end{figure}

@@

[NOTE: I SUGGEST TO:

1. SHOW THE MOMENT 0 AND 1 ONLY MAPS OF C18O, OCS, H2CO AND HCO+ + SPECTRA TOWARD THE CENTER AND +/- 30 AU (FOR EACH SPECIES, YOU CAN DO A 3-PANEL FIGURE SHOWING THE TWO MOMENTS AND A PANEL WITH THE 3 SPECTRA AT DIFFERENT LOCATIONS)
+  THE CONTINUUM MAP 
3. SHOW THE SIMULATED ALMA OBSERVATIONS (SAME STYLE AS BEFORE) AND ALSO SHOW A SEPARATE FIGURE WITH THE PV DIAGRAMS OF ALL SPECIES. 
]

[NOTE: THE FOLLOWING PARAGRAPH IS NOT CLEAR AT ALL !!! PLEASE RE-WRITE MAKING SURE THAT THE SENTENCES MAKE SENSE AND THERE ARE NO TYPOS.  PLEASE MAKE THE FIGURES AS SUGGESTED AND THEN RE-WRITE THIS PART.]
%Simulated molecules were confined to speices where the chemical abundance in the disc and envelope are known, or have simulated abundances. In addition radative transfer simulations are limited to are limited to molecules where the collisional and Einstein co-efficients are known. Lines for these molecules were simulated with LIME for freqencies 50$\sim$500$\,$GHz, with focus on lines which could be observed in ALMA band 7 and molecules which are not typically seen in outflows, which havent been included in these simulations and could potentially contaminate disc observations. The results presented in this section represent those lines which show dectectable emission/absorption which can be used to trace either spiral structure or rotation.\newline

%Figures showing the RT results in a few molecules/transitions (CO, HCO$^+$, HCN, OCS, H2CO, NH3 --- maybe H2O,H2$^{18}$O [to try first])
The results of these simulations for the molecules OCS, HCO$^+$, H$_2$CO and isotopologues of CO are shown here, other molecules simulated but not shown include HCN, HNC, HNO, HCS$^+$ and  CS.
For the purpose of simulating observations the model was placed at roughly the distance of nearby low-mass star forming regions (100$\,$pc) and is inclined at 30$^\circ$ to edge on. From these observations integrated intensity, intensity weighted velocity and position velocity diagrams through the centre of the model were created.
The simulations done are focused upon frequencies within ALMA band 7, selected to give the best trade off between resolution and sensitivity for early ALMA science {\bf IS BAND 7 STILL THE BEST FOR US TO LOOK AT SEEING AS WE ARE NOW LOOKING AT ALMA FULL CAPABILITY RATHER THAN CYCLE1?}.
(Note the moment 1 and 0 maps were created by integrating between -12.5 to -0.5 km$\,$s$^{-1}$ and +0.5 to +12.5 km$\,$s$^{-1}$ to avoid being dominated by the contribution from the envelope, this can be seen in some PV diagrams as the strong absorption feature at all positions around zero velocity, moment 1 maps are shown with a cutoff of 1/100 of the peak emission/absorption value)\newline
%
%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageCO_3-2_30deg_contSub.png}
%
% \caption{CO J=3-2 Continuum subtracted mom0}
%\end{figure}
%
%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageCO_3-2_30deg_mom1.png}
%
% \caption{CO J=3-2 moment 1 map}
 %\end{figure}
%
%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageCO_3-2_30deg_PV_centre.png}
%
% \caption{CO J=3-2 PV through y=0''}
%\end{figure}

%Figures 6,7 and 8 show synthetic images of the CO J=3-2 line at 345.8$\,$GHz. As the upper level for this line is at 33$\,$K above the ground state this transition should be excited through out disc. given the large and uniform nature of the abundance of CO in the disc this line will be optically thick and this is only mapping the outer regions of the disc. In common with most transitions simulated, the line is seen almost exclusively in absorption against the disc continuum. The region of emission in the centre is where there is no continuum background to be absorbed.  Some indications of spiral structure can be seen but are indistinct. {\bf not sure how useful these are, might be worth cutting them?}\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/continuum.png}
 \label{points}
 \caption{A 300$\,$GHz continuum image of the model, created from the average of ten LIME runs.}
\end{figure}

\begin{figure*}
 \includegraphics[width=168mm]{Figures/sim/imageOCS_28-27_30deg_composite_all.png}
 \label{OCS_all}
 \caption{OCS 28-27 {\bf Left:} Continuum subtracted integrated intensity map, {\bf Centre:} Continuum subtracted spectra from Y=0'', X=+0.1'' (bottom), X=-0.1'' (centre) and X=0'' (top) spaced by 10$\,$K for clarity, {\bf Right:} Intensity weighted velocity map}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageOCS_28-27_30deg_composite_PV_centre.png}

@@

\begin{figure*}
 \includegraphics[width=168mm]{Figures/sim/imageH2CO_4-0-4--3-0-3_composite_all.png}
 \label{h2co_all}
 \caption{H$_2$CO 4$_{0\,4}$ - 3$_{\,0\,3}$ {\bf Left:} Continuum subtracted integrated intensity map, {\bf Centre:} Continuum subtracted spectra from Y=0'', X=+0.1'' (bottom), X=-0.1'' (centre) and X=0'' (top) spaced by 50$\,$K for clarity, {\bf Right:} Intensity weighted velocity map}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageH2CO_4-0-4->3-0-3_30deg_PV_centre.png}

@@

\begin{figure*}
 \includegraphics[width=168mm]{Figures/sim/imageHCOp_1-0_30deg_composite_all.png}
 \label{h2co_all}
 \caption{HCO$^+$ 1-0 {\bf Left:} Continuum subtracted integrated intensity map, {\bf Centre:} Continuum subtracted spectra from Y=0'', X=+0.4'' (bottom), X=-0.4'' (centre) and X=0'' (top) spaced by 40$\,$K for clarity, {\bf Right:} Intensity weighted velocity map}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageHCOp_1-0_30deg_PV_centre.png}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/casa_cont_337GHz.png}

@@

\begin{figure}
 \includegraphics[width=54mm]{Figures/sim/casa_imageOCS_28-27_30deg_composite_ALMAwidth_big_dirty_PV_centre.png}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/casa_imageOCS_28-27_30deg_composite_ALMAwidth_big_dirty_contSub.png}

@@

\begin{figure}
 \includegraphics[width=54mm]{Figures/sim/casa_imageC17O_3-2_30deg_composite_ALMAwidth_dirty_PV_centre.png}

 \caption{C17O 3-2 30 deg PV through centre simulated for ALMA most extended configuration}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/casa_C17O_3-2_mom0.png}

 \caption{C17O 3-2 30 deg integrated intensity simulated for ALMA most extended configuration}
\end{figure}

\begin{figure}
 \includegraphics[width=63mm]{Figures/sim/casa_imageHCOp_1-0_30deg_composite_ALMAwidth_dirty_PV_centre.png}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/casa_imageHCOp_1-0_30deg_composite_ALMAwidth_dirty_contSub.png}

@@

@@

In this paper we have performed radiative transfer simulations of a hybrid model comprising a 0.4$\, M_\odot$ self gravitating disc with radius 64$\,AU$ showing spiral density waves, surrounded by an envelope simulated as a collapsing 10$\,M_\odot$ BE-sphere. CASA simulations show that at a distance of 100$\,$pc both spatial resolution of the spirals in such a disc and extraction of kinematic and temperature from molecular lines are possible in ALMA band 7. Our simulations show that many  molecular species are predominantly seen in absorption towards the centre of  self-gravitating protoplanetary discs. The quiescent nature of the envelope around such discs only obscures them within $\pm\,0.5 km\,s^{-1}$.\newline
One assumption made in this model is that the gas and dust are in thermal equilibrium.If they are not and the dust is significantly cooler than the gas then transitions may not show up in absorption.\newline
%The mass of the disc used (0.4$\,\rm{M}_\odot$) is on the high end for early discs (citations for estimate of early class 1 disc masses).\newline
In the model spirals shock heat the mid-plane layer and the colder more diffuse gas absorbs the continuum from it, this results in molecular lines being seen in absorbtion whenever lines of sight pass through the cold diffuse gas at larger heights on to the hot dense midplane of the disc. Rotation curves can be gathered from these observations even with spiral structure\newline
Outflows could contaminate measurements of rotation curves meaning that speices  which commonly trace outflows, such as CO and HCO$^+$ are not going to be good tracers of disc rotation. However speices such as OCS and C$^{18}$O which are not commonly seen in outflows (Yildiz et al 2012,)can be used to trace disc rotation.

@@

\section{Grid Construction} \label{sec:gridding} 
In order to construct the grid, candidate points are randomly selected from the volume to be simulated. These candidates then have their density and molecular density compared against a reference point in order to decide if the point is to be used in the grid or not. [NOTE: THIS FIRST SENTENCE DOES NOT MAKE SENSE. PLEASE RE-WRITE IT AND MAKE SURE THAT IT IS CLEAR AND THERE ARE NO TYPOS.] Candidate grid points are selected at random in cylindrical co-ordinates, linearly spaced in z and $\phi$ and logarithmically spaced in r. For each point to be selected, a random number $\alpha$ is drawn from the semi-open set [0,$\,$1) as a threshold. After selection of random co-ordinates, the hydrogen density and molecular density at the candidate point (n and m, respectively) are compared against the densities of a reference point on the inner edge of the disc (n$_0$ and m$_0$). If $\alpha<\left( \frac{n}{n_0} \right)^{0.3}$ or $\alpha< \left( \frac{m}{m_0} \right)^{0.3}$ then the point is selected for use, otherwise another r, $\phi$, z co-ordinate is selected and this becomes the candidate point. The function comparing the candidate point to the reference point and the candidate point distribution were selected empirically to sample all the scales while ensuring that the majority of points are located in the inner disc where the density is higher. 20\% of these points are forced to be at radii greater than $\sqrt{R_{min}R_{max}}$ (where $R_{min}$ and $R_{max}$ are the inner and outer radius of the model) in order to stop too many of the selecected points clustering in the high density disc and leaving the envelope undersampled. In addiation to this method of selection, 5\% of the points are linearly distributed in x, y and z with no bias with regards to density or abundance. This provides a minimum level of sampling for the large low density regions in the outer parts of the simulated volume. See figure \ref{points} for an example of the points distribution in r, z. \newline

@@

Figures showing different inclinations  
\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_15deg_contSub.png}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_15deg_mom1.png}
%
% \caption{C18O 3-2 15 deg mom1map}
%\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_15deg_PV_centre.png}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_30deg_contSub.png}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_30deg_mom1.png}
%
% \caption{C18O 3-2 30 deg mom1map}
%\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_30deg_PV_centre.png}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_45deg_contSub.png}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_45deg_mom1.png}
%
% \caption{C18O 3-2 45 deg mom1map}
%\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_45deg_PV_centre.png}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_60deg_contSub.png}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_60deg_mom1.png}
%
% \caption{C18O 3-2 60 deg mom1map}
%\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_60deg_PV_centre.png}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_75deg_contSub.png}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_75deg_mom1.png}
%
% \caption{C18O 3-2 75 deg mom1map}
%\end{figure}

\begin{figure}
  \showthe\columnwidth % Use this to determine the width of the figure.
  \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_75deg_PV_centre.png}
  \caption{C18O 3-2 75 deg PV through centre}
\end{figure}

>>>>

\title[Young protoplanetary discs]{Simulated observations of young protoplanetary discs}
\author[Tom  Douglas, Paola Caselli, Et al.]{Tom Douglas$^{1}$\thanks{E-mail:
pytd@leeds.ac.uk}, Paola Caselli$^{1}$, Aaron Boley$^{2}$, Richard Durisen$^{3}$, Tom Hartquist$^{1}$, John Ilee$^{1}$, Jonathon Rawlings$^{4}$ \\
$^{1}$School of Physics and Astronomy, University of Leeds, Leeds LS2 9JT, UK \\
$^{2}$Department of Astronomy, University of Florida, 211 Bryant Space Center, PO Box 112055, USA\\
$^{3}$Department of Astronomy, Indiana University, 727 East 3rd Street, Swain West 319, Bloomington, IN 47405, USA\\
%$^{4}$Department of Physics & Astronomy, University College London, London WC1E 6BT, UK\\
}
\begin{document}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/L1544model_used_legend.eps}
 \label{fig:l1544_model}
 \caption{Model of the pre-stellar core L1544 used as the envelope of the young protoplanetary disc in the hybrid model. Showing temperature (red) and dust (green) temperature in kelvin, log number density (blue) in cm$^{-3}$ and inward velocity$\times$100 (cyan) in m$\,$s$^{-1}$. Adapted from Keto \& Caselli (2010) and Keto et al. (2012, in preparation).}
\end{figure}

@@

\begin{figure*}
 \includegraphics[width=168mm]{Figures/model/columnDensities2.eps}
 \label{Chemistry} 
 \caption{Column densities of OCS, H$_2$CO and HCO$^+$ used in the disc model. Figure adapted from Ilee at al. (2011) [SHOULD WE ALSO ADD C18O COLUMN DENSITY AS WE WANT TO DISCUSS THAT LATER ?? ]}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/rotational_velocities.eps}
 \label{velocity}
 \caption{Azimuthaly averaged rotational velocity in the disc mid-plane with best fit curve}
\end{figure}

@@

\subsection{The radiative transfer code} \label{subsec:radiative_transfer_code}
%-describe the RT used (LIME) 
The radiative transfer program used, LIME (Brinch \&Hogerheijde 2010), calculates line intensities based on a weighted sample of randomly chosen points in a continuous 3D model. The method of selecting these points is given in the grid-construction section. At each of these points, the density of the main collision partner (H$_2$), gas and dust temperatures, velocity, molecular abundances and turbulent velocity are taken from the model. These points are then smoothed by Lloyds algorithm (Lloyd 1982) in order to minimise the variation in distance between points whilst keeping the same underlying distribution. These points are then connected by Delaunay triangulation \footnote{In three dimensions this means that if four points are conneted into a tetrahedron, the sphere circumscribing these four points contains no other points. It can be shown that this connection is unique for a given set of points.} and it is down these paths that photon propagation is restricted (figures \ref{grid}). The level populations of the selected molecules are calculated at each of these points from collisional and radiative (de)excitation and the local radiation field is calculated. This is repeated 20 times with the populations of each level converging towards a single value. This number of iterations is sufficient for the signal to noise ratio of the level populations (as defined in Brinch \&Hogerheijde 2010) to exceed 1000 in 99\% of the points, ensuring that the simulation has converged on a stable level population. \newline
After 20 iterations of updateing the level populations the model is ray-traced in order to produce synthetic brighness maps. In order to minimise the ``blockyness'' of the output images, resulting from the grid construction, the average of ten seperate runs was taken (figure \ref{averages}).

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/average_comparison.eps}
 \label{averages}
 \caption{A plot showing the difference between a single LIME continuum image at 300$\,$GHz (right) and the average of ten such images from different runs of the same model in LIME (left).}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/Lime_grid3.eps}
 \label{grid}
 \caption{A plot of the points selected by the griding process and the paths down which photons can propagate overlaid on a smoothed density model. The points are more concentrated at small radii and in the densest regions.}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/lime_points_rz_histo.eps}
 \label{points}
 \caption{A 2D histogram of the point distribution throughout the model. The disc and envelope can be seen as two separate entities which have to be sampled using different point distributions}
\end{figure}

@@

Simulations are limited to molecules for which we have abundances in both the disc and envelope, and also have calculated Einstein and collisional co-efficients. The Molecular data used is from the lamda database (reference? http://home.strw.leidenuniv.nl/$\sim$moldata/). Lines for the molecules which satisfy these conditions were simulated with LIME for freqencies between 50$\sim$500$\,$GHz, with a focus on lines which could be observed in ALMA band 7 and molecules which are not typically seen in outflows. The results presented in this section represent those lines which show dectectable emission/absorption which can be used to trace either spiral structure or rotation.\newline

Transitions focused upon are CO 3-2 (and isotopologues) at 345.8GHz with an upper energy level of 33.2K, HCO+ 1-0 at 89.2GHZ and 4.3K, OCS 28-27 at 340.4GHz and 237.0K and H2CO $4_{04}$-3$_{03}$ at 290.6GHz and 34.9K {\bf maybe worth including a higher frequency line as well? maybe CS at ~650GHz}. Other molecules simulated but not shown include HCN, HNC, HNO, HCS$^+$ and CS.\newline

For the purpose of simulating observations the model was placed at roughly the distance of nearby low-mass star forming regions (100$\,$pc) and is inclined at 30$^\circ$ to edge on. From these observations integrated intensity, intensity weighted velocity and position velocity diagrams through the centre of the model were created.
The simulations done are focused upon frequencies within ALMA band 7, selected to give the best trade off between resolution and sensitivity for early ALMA science.
(Note the moment 1 and 0 maps were created by integrating between -12.5 to -0.5 km$\,$s$^{-1}$ and +0.5 to +12.5 km$\,$s$^{-1}$ to avoid being dominated by the contribution from the envelope, this can be seen in some PV diagrams as the strong absorption feature at all positions around zero velocity, moment 1 maps are shown with a cutoff of 1/100 of the peak emission/absorption value)\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/continuum.eps}
 \label{points}
 \caption{A 300$\,$GHz continuum image of the model, created from the average of ten LIME runs.}
\end{figure}

\begin{figure*}
 \includegraphics[width=168mm]{Figures/sim/imageOCS_28-27_30deg_composite_all.eps}
 \label{OCS_all}
 \caption{OCS 28-27 {\bf Left:} Continuum subtracted integrated intensity map, {\bf Right:} Intensity weighted velocity map}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageOCS_28-27_30deg_composite_PV_centre.eps}

@@

\begin{figure*}
 \includegraphics[width=168mm]{Figures/sim/imageH2CO_4-0-4--3-0-3_30deg_composite_all.eps}
 \label{h2co_all}
 \caption{H$_2$CO 4$_{0\,4}$ - 3$_{\,0\,3}$ {\bf Left:} Continuum subtracted integrated intensity map, {\bf Right:} Intensity weighted velocity map}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageH2CO_4-0-4--3-0-3_30deg_composite_PV_centre.eps}

@@

\begin{figure*}
 \includegraphics[width=168mm]{Figures/sim/imageHCOp_1-0_30deg_composite_all.eps}
 \label{h2co_all}
 \caption{HCO$^+$ 1-0 {\bf Left:} Continuum subtracted integrated intensity map, {\bf Right:} Intensity weighted velocity map}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageHCOp_1-0_30deg_composite_PV_centre.eps}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/casa_cont_337GHz.eps}

@@

\begin{figure}
 \includegraphics[width=54mm]{Figures/sim/casa_imageOCS_28-27_30deg_composite_ALMAwidth_big_dirty_PV_centre.eps}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/casa_imageOCS_28-27_30deg_composite_ALMAwidth_big_dirty_contSub.eps}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/casa_HCOp_spectra_40_AUsplit.eps}

 \caption{HCO$^+$ 1-0 30 deg spectra at Y=0, X=+40AU (bottom), 0AU (middle) and -40AU (Top)}
\end{figure}

\begin{figure}
 \includegraphics[width=54mm]{Figures/sim/casa_imageC17O_3-2_30deg_composite_ALMAwidth_dirty_PV_centre.eps}

 \caption{C$^{17}$O 3-2 30 deg PV through centre simulated for ALMA most extended configuration}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/casa_C17O_3-2_mom0.eps}

 \caption{C$^{17}$O 3-2 30 deg integrated intensity simulated for ALMA most extended configuration}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/casa_C17O_spectra_20AUsplit.eps}

 \caption{C$^{17}$O 1-0 30 deg spectra at Y=0, X=+20AU (bottom), 0AU (middle) and -20AU (Top)}
\end{figure}

\begin{figure}
 \includegraphics[width=63mm]{Figures/sim/casa_imageHCOp_1-0_30deg_composite_ALMAwidth_dirty_PV_centre.eps}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/casa_imageHCOp_1-0_30deg_composite_ALMAwidth_dirty_contSub.eps}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/casa_OCS_spectra_10AUsplit.eps}

 \caption{OCS 28-27 30 deg spectra at Y=0, X=+10AU (bottom), 0AU (middle) and -10AU (Top)}
\end{figure}

@@

In this paper we have performed radiative transfer simulations of a hybrid model comprising a 0.4$\, M_\odot$ self gravitating disc with radius 64$\,AU$ showing spiral density waves, surrounded by an envelope simulated as a collapsing 10$\,M_\odot$ BE-sphere. CASA simulations show that at a distance of 100$\,$pc both spatial resolution of the spirals in such a disc and extraction of kinematic and temperature from molecular lines are possible in ALMA band 7. Our simulations show that many  molecular species are predominantly seen in absorption towards the centre of  self-gravitating protoplanetary discs. The quiescent nature of the envelope around such discs only obscures them within $\pm\,0.5 km\,s^{-1}$.\newline
One assumption made in this model is that the gas and dust are in thermal equilibrium.If they are not and the dust is significantly cooler than the gas then transitions may not show up in absorption.\newline
%The mass of the disc used (0.4$\,\rm{M}_\odot$) is on the high end for early discs (citations for estimate of early class 1 disc masses).\newline
In the model spirals shock heat the mid-plane layer and the colder more diffuse gas absorbs the continuum from it, this results in molecular lines being seen in absorbtion whenever lines of sight pass through the cold diffuse gas at larger heights on to the hot dense midplane of the disc. Rotation curves can be gathered from these observations even with spiral structure\newline
Outflows could contaminate measurements of rotation curves meaning that speices  which commonly trace outflows, such as CO and HCO$^+$ are not going to be good tracers of disc rotation. However speices such as OCS and C$^{18}$O which are not commonly seen in outflows (e.g. van der Tak et al. 2003, Stanke et al. 2007, Yildiz et al 2012, Ren et al. 2011)can be used to trace disc rotation.

@@

\section{Grid Construction} \label{sec:gridding} 
In order to construct the grid, candidate points are randomly selected from the volume to be simulated. These candidates then have their density and molecular density compared against a reference point in order to decide if the point is to be used in the grid or not. Candidate grid points are selected at random in cylindrical co-ordinates, linearly spaced in z and $\phi$ and logarithmically spaced in r. For each point to be selected, a random number $\alpha$ is drawn from the semi-open set [0,$\,$1) as a threshold. After selection of random co-ordinates, the hydrogen density and molecular density at the candidate point (n and m, respectively) are compared against the densities of a reference point on the inner edge of the disc (n$_0$ and m$_0$). If $\alpha<\left( \frac{n}{n_0} \right)^{0.3}$ or $\alpha< \left( \frac{m}{m_0} \right)^{0.3}$ then the point is selected for use, otherwise another r, $\phi$, z co-ordinate is selected and this becomes the candidate point. The function comparing the candidate point to the reference point and the candidate point distribution were selected empirically to sample all the scales while ensuring that the majority of points are located in the inner disc where the density is higher. 20\% of these points are forced to be at radii greater than $\sqrt{R_{min}R_{max}}$ (where $R_{min}$ and $R_{max}$ are the inner and outer radius of the model) in order to stop too many of the selecected points clustering in the high density disc and leaving the envelope undersampled. In addiation to this method of selection, 5\% of the points are linearly distributed in x, y and z with no bias with regards to density or abundance. This provides a minimum level of sampling for the large low density regions in the outer parts of the simulated volume. See figure \ref{points} for an example of the points distribution in r, z. \newline

@@

Figures showing different inclinations  
\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_15deg_contSub.eps}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_15deg_mom1.eps}
%
% \caption{C18O 3-2 15 deg mom1map}
%\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_15deg_PV_centre.eps}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_30deg_contSub.eps}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_30deg_mom1.eps}
%
% \caption{C18O 3-2 30 deg mom1map}
%\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_30deg_PV_centre.eps}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_45deg_contSub.eps}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_45deg_mom1.eps}
%
% \caption{C18O 3-2 45 deg mom1map}
%\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_45deg_PV_centre.eps}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_60deg_contSub.eps}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_60deg_mom1.eps}
%
% \caption{C18O 3-2 60 deg mom1map}
%\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_60deg_PV_centre.eps}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_75deg_contSub.eps}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_75deg_mom1.eps}
%
% \caption{C18O 3-2 75 deg mom1map}
%\end{figure}

\begin{figure}
%  \showthe\columnwidth % Use this to determine the width of the figure.
  \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_75deg_PV_centre.eps}
  \caption{C18O 3-2 75 deg PV through centre}
\end{figure}
